Hector Permafrost

### 2020-02-06 analysis/paper/paper.Rmd
Revise permafrost methods for new Hector changes
- Use lognormal CDF
- RH{CH4} is on top of regular RH, not a trade-off

---
title: "Hector Permafrost"
author:
  - Alexey N. Shiklomanov
  - Ben Bond-Lamberty
  - Ben Kravitz
  - Corinne Hartin
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx"
    rmarkdown::github_document:
      html_preview: false
bibliography: ["references.bib", "packages.bib"]
csl: "../templates/biogeosciences.csl"
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

To a first approximation, $\Phi[t]$ is a function of air temperature ($T[t]_{air}$).
[Kessler (2015)][kessler] assume this relationship is linear.
However, because the permafrost area fraction is, by definition, bounded by zero (and 1 [^1]), and because deeper permafrost thaws more slowly than shallow permafrost, we use a negative logistic relationship instead:

@@

\[
\Phi[t] = 1 - \left( 1 + a \exp\left( b * T_{air}[t] \right) \right) ^ c
\]

where $a$, $b$, and $c$ are model parameters whose values are described in the _Configuration_ section below.
The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:

\[
\Delta \Phi[t] = \max(\Phi[t] - \Phi[t-1], 0)
\]

In other words, permafrost thaw is permanent -- once it thaws, it does not re-freeze, even if the temperature drops.

@@

\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {T}[i,t] / 10} \]

@@

In the original formulation, all of the total heterotrophic respiration was assumed to be CO~2~ and natural CH~4~ emissions were entirely separate.
In the revised version here, total heterotrophic respiration is divided into CO~2~ ($RH_{CO_2}[i,t]$) and CH~4~ ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):

@@

\[ RH_{CO_2}[i,t] = (1 - f_{CH_4}[i]) RH[i,t] \]

>>>>

---
title: "Hector Permafrost"
author:
  - Alexey N. Shiklomanov
  - Ben Bond-Lamberty
  - Ben Kravitz
  - Corinne Hartin
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::pdf_document2:
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx"
    rmarkdown::github_document:
      html_preview: false
bibliography: ["references.bib", "packages.bib"]
csl: "../templates/biogeosciences.csl"
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

To a first approximation, $\Phi[t]$ is a function of air temperature ($T[t]_{air}$).
[Kessler (2015)][kessler] assume this relationship is linear.
However, because the permafrost area fraction is, by definition, bounded by zero (and 1 [^1]), and because deeper permafrost thaws more slowly than shallow permafrost, we use the lognormal cumulative distribution function ($NCDF(\log(x) | \mu, \sigma)$) instead:

@@

\[
\Phi[t] = 1 - NCDF(\log(x) | \mu, \sigma)
\]

where $\mu$ and $\sigma$ are model parameters whose values are described in the _Configuration_ section below.
The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:

\[
\Delta \Phi[t] = \Phi[t] - \Phi[t-1]
\]

@@

\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {{T}[i,t] / 10} \]

@@

In the original formulation, natural CH~4~ emissions were a prescribed constant value.
In the revised version here, natural CH~4~ emissions are a fraction of heterotrophic respiration ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):

@@
